---
title: "COVID positive rate estimation using social media- Tarrant county, Tx, USA"
output:
  html_document:
    df_print: paged
---



#### Loading libraries and functions

```{r message=FALSE, warning=FALSE}

setwd("C:/Users/mfondoum/Dropbox/COVID")

source("libraries.R")
### sourcing function covidcast
#source("Api.R")
#source("covid_func.R")

source("loading-and-testing-the-data.R")

```


### Getting census data per location (Function sourced from covid-func.R)

```{r}

#### get_census_data (Function sourced from covid-func.R)

#dfw_data <- get_census_data()

```


### Getting Tarrant Dataset

```{r}

### Secondary data source gathering (Function sourced from covid-func.R)
#covid_data_load()

shift_days <- 90

data_60 <- covid_data_60()

tarrant_data_ma <- get_tarrant_df()

#### Splitting into train test

train_index <- index(tarrant_data_ma)[1:round(nrow(tarrant_data_ma)*0.75)]

train_data <- tarrant_data_ma[train_index,]

test_data <- tarrant_data_ma[-train_index,]

nrow(test_data)+nrow(train_data) == nrow(tarrant_data_ma)

```



#### Building model

```{r}

lm_t1 <- fit_model(df=na.omit(train_data))

summary(lm_t1)

```


#### Predicting Tarrant County

```{r}

tarrant_pred_train <- pred_tarrant_model(model = lm_t1, df=na.omit(train_data))

tarrant_pred_test <- predict(lm_t1, na.omit(test_data[,3:6]))

### Combining test data Actual vs Prediction 
tarrant_test_actual <- test_data[which(rownames(test_data) %in% names(tarrant_pred_test)),c("date", "positivity_rate_7d")] %>% rename(Actual=positivity_rate_7d)

tarrant_pred_test <- tibble(tarrant_test_actual, Fitted=tarrant_pred_test)

### Combining Train and Test Predictions results

tarrant_pred <- bind_rows(na.omit(tarrant_pred_train), tarrant_pred_test)


#tarrant_fitted <- c(lm_t1$fitted.values, tarrant_pred_test)

#tarrant_actual <- rbind(train_data, na.omit(test_data))[as.numeric(names(tarrant_fitted)),]

#tarrant_final <- cbind(tarrant_actual, Fitted=tarrant_fitted)

#names(tarrant_final)[7]<- "Actual"

#tarrant_pred <- tarrant_final

test_accuracy <- tarrant_pred %>% dplyr::filter(between(date, as.Date("2020-08-25"), as.Date("2020-09-08"))) %>% dplyr::mutate(Res_sq=(Actual-Fitted)^2, Tot_seq=(Actual-mean(Actual))^2)



### Calculating the test accuracy based on the diag table (this is good for binary prediction or multinomila prediction)

tab <- table(test_accuracy$Actual, test_accuracy$Fitted)

#tab

1-sum(diag(tab))/sum(tab)

test_accuracy = sum(diag(tab))/sum(tab)

test_accuracy

### Train accuracy (Accuracy on train proves that this accuracy calculationis not suitable for continuous data)

tab1 <- table(na.omit(tarrant_pred_train)$Fitted, na.omit(tarrant_pred_train)$Actual)

train_accuracy = sum(diag(tab1))/sum(tab1)

train_accuracy


```


#### Plotting Tarrant County Prediction

```{r}

colors <- c("Actual"="steelblue", "Fitted"="#E69F00")
  
p <- ggplot(data = tarrant_pred, aes(x=date))+
    geom_point(aes(y=Actual), color="steelblue")+
    geom_line(aes(y=Actual), color="steelblue")+
    geom_point(aes(y=Fitted), color="#E69F00")+
    geom_line(aes(y=Fitted), color="#E69F00")+
    scale_x_date(date_breaks = "weeks")+
    geom_vline(xintercept = as.numeric(tarrant_pred[tarrant_pred$date==as.Date("2020-08-25"),"date"]),color="red", linetype="dashed")+
    annotate(geom = "segment", x=c(as.Date("2020-09-02"), as.Date("2020-09-02")), xend = c(as.Date("2020-09-04"),as.Date("2020-09-04")), y=c(19, 18), yend = c(19, 18), color=c("steelblue", "#E69F00"), size=1)+
    annotate(geom = "text", x=c(as.Date("2020-08-13"),as.Date("2020-08-21"), as.Date("2020-09-07"), as.Date("2020-09-07"),as.Date("2020-08-13"),as.Date("2020-08-30")), y=c(19.5, 19.5, 19, 18,16,16), label=c("R-Squared:",round(summary(lm_t1)$r.squared,2), "Actual", "Fitted","Train data", "Test data"), size=4, color="blue")+
    theme_bw()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.subtitle = element_text(color="blue"), plot.title = element_text(size=12,colour = "black", face = "bold"))+
    labs(title = "Tarrant County daily risk",subtitle = "(MyPHI Covid Analysis)", x="Date", y="Positivity rates")+
    scale_color_manual(name="Legend", values = colors)

ggplotly(p)

```

#### Visualizing the data

```{r}

plot(lm_t1)

```


```{r}

ggplot(data = tarrant_pred, aes(x=Fitted, y=Actual))+geom_point()+geom_smooth(method = "lm")

summary(lm_t1)

```



#### DFW counties Risk map

```{r}

#dfw_fips <- c(48085,48113, 48121, 48139, 48221, 48231, 48251, 48257, 48367, 48397, 48425, 48439, 48497)

#dfw_counties <- tibble(county=fips_to_name(dfw_fips))


get_county_last_pred <- function(county1){

  test_ma <- get_county_test(county_name = county1)

  test_pred <- pred_county(model = lm_t1, df=test_ma)
  
  test_pred <- na.omit(test_pred)
  
  as.numeric(test_pred$Fitted[length(test_pred$Fitted)])

}

all_counties_pred <- lapply(dfw_counties$county, function(x) get_county_last_pred(county=x))

all_counties_pred <- Map(get_county_last_pred, dfw_counties$county)

all_counties_pred <-as.data.frame(unlist(all_counties_pred))

names(all_counties_pred)<-c("Risk")
all_counties_pred$county <- fips_to_name(rownames(all_counties_pred))

all_counties_pred <- full_join(dfw_counties, all_counties_pred, by="county")

all_counties_pred$Risk <- round(all_counties_pred$Risk,2)

all_counties_pred

```


### plotting DFW Counties Risk on map

```{r}

library(usmap)
#dfw_fips <- c(48085,48113, 48121, 48139, 48221, 48231, 48251, 48257, 48367, 48397, 48425, 48439, 48497)

#dfw_counties <- fips_to_name(dfw_fips)

all_counties_pred <- cbind(all_counties_pred,fips=dfw_fips)

#all_counties_pred$fips <- name_to_fips(all_counties_pred$county)

p <- plot_usmap("counties",color = "red", values="Risk", alpha=0.25,include = c(dfw_fips), data = all_counties_pred, labels = TRUE) +
  scale_fill_continuous(name = "Covid risk rate per county", label = scales::comma, low="white", high="red") + 
  theme(legend.position = "right", title = element_text(size = 12, colour = "black", face="bold"))+
  labs(title = "DFW County covid risk prediction. Sep 22/09/2020",
       caption = "These are the states in the Pacific Timezone.")

ggplotly(p)

```


Zipcode level data

```{r}
library(tigris)

us_states <- states()

us_states_20m <- states(cb = TRUE, resolution = "20m")

tx_20m <- us_states_20m[us_states_20m$NAME=="Texas","STATEFP"]

tx_road <- roads(state = "Texas", county = "Tarrant County")

plot(tx_road)


```

```{r}

library(mapproj)
library(maps)

map("county", state.abb="TX" ,projection = "lambert", 
    parameters = c(lat0 = 20, lat1 = 50), wrap = TRUE)

```

#### With Leaflet
```{r}

Tarrant_county <- tibble(
    address = "5615 Forest Hill Drive, TX 76119",
    title="Beal, Harlean Elementary Schools",
    street_address= "5615 Forest Hill Drive, TX 76119"
  ) %>%
  tidygeocoder::geocode(address, method = "osm")

library(leaflet)

Tarrant_county_map <- leaflet() %>% 
  addTiles() %>%
  addMarkers(data = Tarrant_county) %>%
  addPopups(
    data = Tarrant_county, 
    popup = ~paste0("<b>", title, "</b></br>", street_address)
  )

Tarrant_county_map

```



```{r}

nc <- ggplot(data = nc_merged, aes(fill = r_win)) +
  annotation_map_tile(zoom = 6, type = "osm") + 
  geom_sf(alpha = 0.5) +
  scale_fill_manual(values = c("blue", "red")) + 
  geom_sf_label(aes(label = DISTRICT), fill = "white") + 
  theme_void()
nc
```



Geospatial analysis with sf

```{r}
library(sf)
library(tmap)

zipcode_geo <- st_read("cb_2018_us_zcta510_500k.kml")

plot(zipcode_geo[1])

?sf

```



US Boundaries Package

```{r}

install.packages("tmap")


```








### Regression


```{r}

tarrant_lm <- lm(data = tarrant_data, Percent_positivity_7d_Avg ~ Google_grocery_7d_Avg+Google_park_7d_Avg+fb_cli_perc_7d_Avg)

summary(tarrant_lm)

plot(tarrant_lm)

covid_model <- saveRDS(tarrant_lm, "covid_model")

```


### Getting test data

```{r}

tarrant_test <- covid_get(county_name = "Tarrant", start_date = "2020-09-01", end_date = "2020-09-17")

View(tarrant_data)

tarrant_test_data <- covid_7d_ma(tarrant_test, has_response=TRUE)


test_response <- c(10.14,9.57,9.43,9.71,10.0,8.71,9.57,9.29,10.0,9.71) 

tarrant_resp_test <- tibble(Percent_positivity=test_response[1:8])

tarrant_test_data <- cbind(tarrant_test_data, tarrant_resp_test)


tarrant_test_data


############


tarrant_test <- fb_data_tarrant %>% dplyr::select(time_value, value)

#tarrant_test <- covid_7d_ma(tarrant_test)

names(tarrant_test) <- c("date", "cli_perc")


google_test <- google_data %>% filter(between(date, as.Date("2020-08-01"), as.Date("2020-09-17"))& state=="Texas" & county=="Tarrant County") %>% dplyr::select(date,grocery.and.pharmacy, parks, state, county)

tarrant_fb_test <- tarrant_test %>% filter(between(date, min(google_test$date), max(google_test$date)))

tarrant_fb_test <- tarrant_fb_test %>% dplyr::select(cli_perc)


tarrant_final <- cbind(google_test, tarrant_fb_test)

tarrant_final <- tarrant_final %>% dplyr::select(date, grocery.and.pharmacy, parks, cli_perc)

## Moving Average

tarrant_final_ma <- tarrant_final %>% mutate(Google_park_7d_Avg = round(rollmean(parks,7, fill = NA, align = "right"),1),
             Google_grocery_7d_Avg = round(rollmean(grocery.and.pharmacy, 7, fill = NA, align = "right"),1),
             fb_cli_perc_7d_Avg = round(rollmean(cli_perc*100, 7, fill = NA, align = "right"),1))
             #fb_clic_perc_7d_Avg = round(rollmean(clic_perc, 7, fill = NA, align = "right"),1),
             #ft_workaway_7d_Avg = round(rollmean(ft_workaway, 7, fill = NA, align = "right"),3),
             #Percent_positivity_7d_Avg = round(rollmean(Percent_positivity, 7, fill = NA, align = "right"),2))


tarrant_final_ma

tarrant_pos <- read.csv("tarrant_county_data.csv")

head(tarrant_pos,20)

```


### Predictions

#### Predictions on Tarrant County

```{r}

pred_res <- predict(tarrant_lm, tarrant_test_data)


Actual_train <- tarrant_data %>% filter(between(date, as.Date("2020-08-01"), as.Date("2020-08-31"))) %>% select(date,Actual=Percent_positivity_7d_Avg)

predictor_train <- tarrant_data %>% filter(between(date, as.Date("2020-08-01"), as.Date("2020-08-31"))) %>% select(Google_grocery_7d_Avg, Google_park_7d_Avg, fb_cli_perc_7d_Avg)



pred_train <- predict(tarrant_lm, predictor_train)

pred_out <- tibble(date=tarrant_test_data$date, Fitted=round(pred_res,2), Actual=tarrant_test_data$Percent_positivity)

df_training <- cbind(Actual_train, pred_train)

names(df_training) <- c("date", "Actual", "Fitted")

p <- ggplot()+geom_point(data = df_training, aes(x=date, y=Actual), color="green")+geom_point(data = df_training, aes(x=date, y=Fitted), color="red")


ggplotly(p)

head(df_training)

```

Plotting Prediction results
```{r}
library(plotly)

pp <- ggplot()

pp1 <- geom_point(data = pred_out, aes(x=date, y=Actual), color='darkblue')

pp2 <- geom_point(data = pred_out, aes(x=date, y=Fitted), color='red')

ggplotly(pp+pp1+pp2)


```


#### Predictions on another County


```{r}
### Uncomment the line below to load the savce covid_model
# tarrant_lm <- readRDS("covid_model")

covid_predic(county = "Tarrant", startdate = "2020-05-01", enddate = "2020-09-08")

```






#### Clustering

K-means


```{r}

#kmeans()

```


PCA

```{r}


#library(splines)

#fit_spline <- lm(Percent_positivity_7d_Avg ~ Google_park_7d_Avg)

```


#### Time series

```{r}

library(ggplot2)

library(dplyr)

p <- ggplot()
p0 <- geom_line(data = data_new,aes(x=Day, y=Percent_positivity_7d_Avg), color='green')
p00 <- annotate(geom = 'text', x=last(data_new$Day), y=last(data_new$Percent_positivity_7d_Avg), label='Percent_positivity_7d_Avg')

p1 <- geom_line(data = data_new,aes(x=Day, y=Google_park_7d_Avg), color='red')
p3 <- geom_line(data = data_new,aes(x=Day, y=cli_perc_7d_Avg), color='darkblue')

p2 <- geom_line(data = data_new,aes(x=Day, y=Google_grocery_7d_Avg), color='blue')
p4 <- geom_line(data = data_new,aes(x=Day, y=clic_perc_7d_Avg))
p5 <- geom_line(data = data_new,aes(x=Day, y=ft_workaway_7d_Avg))

```

```{r}

p+p0+p1+p2

p+p0+p3+p4+p5

p+p0+p1+p2+p3+p4+p5

p+p0+p00

```





```{r}
p <- ggplot()
#q0 <- geom_point(data = data_new,aes(x=Day, y=Percent_positivity_7d_Avg), color='green')
q1 <- geom_point(data = data_new,aes(x=Percent_positivity_7d_Avg, y=Google_park_7d_Avg), color='red')
q2 <- geom_point(data = data_new,aes(x=Percent_positivity_7d_Avg, y=Google_grocery_7d_Avg), color='blue')
q3 <- geom_point(data = data_new,aes(x=Percent_positivity_7d_Avg, y=cli_perc_7d_Avg), color='darkblue')
q4 <- geom_point(data = data_new,aes(x=Percent_positivity_7d_Avg, y=clic_perc_7d_Avg))
q5 <- geom_point(data = data_new,aes(x=Percent_positivity_7d_Avg, y=ft_workaway_7d_Avg))

```


```{r}

prediction <- tibble(Day=test_df$Day, Fitted=pred_res)

df_all <- rbind(df,prediction)

actual <- tibble(Actual=c(data_new$Percent_positivity_7d_Avg, test_response))

df_final <- tibble(df_all, actual)

df_final

```

Appended data
```{r}

df1 <- data_new[,c('Google_park_7d_Avg', 'Google_grocery_7d_Avg', 'cli_perc_7d_Avg')]

df2 <- tibble(ght_data_test[,-1],fb_data_val_test)

names(df1) <- c("Park", "Grocery", "cli_perc")

df2 <- df2 %>% dplyr::select(Park,Grocery,cli_perc)

df3 <- add_row(df1, df2)

View(df3)

```



Plotting fitted versus actual

```{r}

qq <- ggplot()

qq1 <- geom_point(data = df_final, aes(x=Day, y=Actual), color='darkblue')

qq2 <- geom_point(data = df_final, aes(x=Day, y=Fitted), color='red')

qq+qq1+qq2+geom_vline(xintercept = as.numeric(df_final$Day[63]))+labs(title = "MyPHI COVID Model Accuracy: Predicted (Red) vs Actual (Blue)")+theme_bw()+theme(title = element_text(colour = "blue"))+annotate(geom = 'text', x=df_final$Day[c(31,67,31)], y=c(18,18,20), label=c("Training", "Test", "R_Sqr = 0.9063"))

```


#### R-squared of test dataset

```{r}

combined_test <- tibble(Percent_positivity_7d_Avg=test_response, test_df)

summary(lm(data = combined_test, Percent_positivity_7d_Avg ~ Google_grocery_7d_Avg+Google_park_7d_Avg+cli_perc_7d_Avg))

```


