---
title: "The COVID-19 Symptom Data Challenge - Phase I Application"
output:
  html_document:
    df_print: paged
  html_notebook:
    code_folding: hide
    css: style.css
---

<p><br><br>

### Supporting schools reopening decision making in Dallas Fort Worth using Facebook symptoms public data

<div>
<br>
<font align="justify">We sought to develop a prediction model that accurately predicts COVID-19 community transmission in a timely manner, to support school-reopening decision making and school-based COVID-19 surveillance strategy. Our model leverages county-level Facebook Symptom Tracker data along with Apple mobility data to predict daily rates of COVID-19 community transmission to guide decision making for safe school reopening, timely closure, or school-based surveillance testing. By using our model, School and Public Health might implement interventions to mitigate the risk for school-based COVID-19 transmission, enhance students’ educational experience, and limit disruptive school closures.</font></div>

</p>

<div>

<div>
<p><br><br>



```{r echo=TRUE}

## Loading parameters
source("libraries.R")
source("params.R")
source("functions.R")

## Getting historical data
source("covid_get_fb.R")
source("covid_get_sg.R")
source("covid_get_apple.R")

# Loading sequentially R codes
source("load-data-ts.R")
source("tf_apl.R")
source("tf_fb_sci.R")
source("tf_sg.R")


```

#### Variables selection 
<p><br>
Loading data from Facebook Symptoms datasets and other public datasets for analysis:

1. Facebook survey data

  * Facebook Covid Like Illness (smoothed and weighted) as "cli"
  * Facebook Community Covid Like Illness (smoothed and weighted) as "cmnt_cli"

2. Safegraph mobility data

  * Average time spent at full-time work location
  * Average time spent at hall-time work location  

3. Apple mobility data

This data contains device change in mobility using Jan 13th, 2020 as reference to track changes.

  * Walking
  * Transit
  * Driving


### Data exploration 

#### Facebook Data {.tabset}

##### Facebook Plot

```{r warning=FALSE, message=FALSE, out.width="80%"}

fb_df_tarrant %>%
  melt(id.vars = c("date"), measure.vars = c("cli", "cmnt_cli")) %>% 
  group_by(variable) %>%
  group_map(~plot_ly(.,x=~date, y=~value,color = ~variable, mode = "lines"), keep=TRUE) %>%
  subplot(nrows = 1, shareX = TRUE, shareY = FALSE) %>%
  layout(title="Facebook CLI vs Community CLI")
  

```

##### Facebook data

```{r warning=FALSE, message=FALSE}

tail(fb_df_tarrant)

```
  


#### Safegraph Data {.tabset}


##### Safegraph Plot

```{r warning=FALSE, message=FALSE,out.width="80%"}

sg_test_ft_ts <- ts(sg_df_tarrant$fulltime, frequency = 7)

sg_test_pt_ts <- ts(sg_df_tarrant$parttime, frequency = 7)

### Autoarima to select the best forecast based on AIC
fit_pt <- auto.arima(sg_test_pt_ts)

fit_ft <- auto.arima(sg_test_ft_ts)

### Forecast length by taking the difference between last day provided by safegraph and today 
len <- as.numeric(today() - last(sg_df_tarrant$date))

## Forecasting 
sg_pt_arima <- forecast(fit_pt, h=len)

sg_ft_arima <- forecast(fit_ft, h=len)

sg_pt_est <- sg_pt_arima$mean[1:len]

sg_ft_est <- sg_ft_arima$mean[1:len]

new_dates <- seq.Date(from=last(sg_df_tarrant$date)+1, to=today(), by="day")

sg_fcst <- tibble(fips=tail(sg_df_tarrant$fips,len), date=new_dates, fulltime=sg_ft_est, parttime=sg_pt_est)

sg_df_tarrant <- bind_rows(sg_df_tarrant, sg_fcst)


sg_df_tarrant %>%
  plot_ly(x=~date) %>%
  add_trace(y=~fulltime, mode="lines", name="Fulltime") %>%
  add_trace(y=~parttime, mode="lines", name="Parttime")


```
##### Safegraph data

```{r warning=FALSE, message=FALSE}

tail(sg_df_tarrant)

```
  



#### Apple Data {.tabset}


##### Apple Plot

```{r warning=FALSE, message=FALSE, out.width="80%"}

apple_mob_tarrant[,1:5] %>%
  plot_ly(x=~date) %>% 
  add_trace(y=~driving, mode="lines", name="driving") %>%
  add_trace(y=~walking, mode="lines", name="walking") %>%
  add_trace(y=~transit, mode="lines", name="transit") %>%
  layout(title="Apple mobility data")

```
##### Mobility data

```{r warning=FALSE, message=FALSE}

tail(apple_mob_tarrant[,1:5])

```


</div>


<div>


# Compiling the final dataset

```{r}

source("final_data.R")

```


#### Features engineering

From apple data, we create a composite indicator by combining the effect of driving and transit variables into a single variable which the transit ratio to transit + driving. $transit/(transit+driving)$. This variable is calculated based on the assumption that the higher the transit, the higher the risk of infection and the higher the driving, the lower the risk of infection, hence driving is at the denominator of the formula and transit as well is there to normalize the value between 0 and 1.

```{r warning=FALSE, message=FALSE}

plot_ly(data = df_final, x=~date) %>% 
  add_trace(y=~comp_indice, mode="lines") %>%
  layout(title="Apple Mobility Composite indicator", yaxis=list(title="Mobility index"))

```

</div>


<div>
### Combined dataset {.tabset}

#### Data

```{r}

df_final

save(df_final, file = "data/df_final.rda")

load("data/df_final.rda")

```

</div>

<div>

```{r}

##############################################################
#                   NEW PHASE                                #
############## Vector Error correction Model #################
require(bvartools)
require(urca)
require(vars)
require(dynlm)
require(forecast)

ndays1 <- 14

k <- nrow(df_final)-ndays1


##############################################################
####  Making the dataset stationary by first differencing ####

df_full_ts <- ts(df_final[,-1])

df_full_ts_sta <- diff(df_full_ts,1)


##############################################################
############## Creating the training and test timeseries ##################
df_final_ts <- df_full_ts[1:k,]
  
df_test <- df_full_ts[(k+1):nrow(df_full_ts),]

df_test_ts <- ts(df_test[,-c(1:2)])

### Stationarizing the test dataset
df_test_ts_sta <- diff(df_test_ts,1)

save(df_test, file="data/df_test.rda")

save(df_test_ts, file="data/df_test_ts.rda")

```

#### Original timeseries

```{r}

df_1 <- df_final[1:k,-1]

names(df_1) <- c("Positivity", "CLI", "Community_CLI", "Fulltime", "Partime", "Apple_comp_index")

plot(ts(df_1), main="Figure 1: Time series plot non-stationary")


```


#### Stationary timeseries
This is the representation of the time series after applying the first difference with the first lag value


```{r}

df_2 <- df_final[1:k,-1]

names(df_2) <- c("Positivity", "CLI", "Community_CLI", "Fulltime", "Partime", "Apple_comp_index")

df_2 <- na.omit(diff(ts(df_2), 1))

plot(df_2, main="Figure 1: Time series plot non-stationary")

```
</div>

### Building the VAR Models {.tabset}

#### VAR Model summary

```{r}

source("functions.R")

##############################################################
################## VAR MODEL BUILDING ########################

#Lag selection
var_lag <- VARselect(df_final_ts, lag.max = 8, type = "both", season = 7)

# Extracting the lag seleciton by AIC
max_lag = as.numeric(var_lag$selection[1])

## Option 1: VAR Model
VAR_pos <- VAR(df_final_ts, lag.max = max_lag, type = "both", season = 7)

summary(VAR_pos$varresult$pos_ma)


```

#### Coefficient testing

```{r}

coeftest(VAR_pos$varresult$pos_ma)

```


### Building the VECM Models {.tabset}

#### VECM Model summary

The model shows 5 point of cointegration by the values of r based on the Johanssen Cointegration testing procedure

```{r}

##############################################################
################## VECM MODEL BUILDING #######################

k_order <- VAR_pos$p

vecm_model1 <- ca.jo(df_final_ts, ecdet = "none", type = "trace", K=k_order, spec = "transitory")

vecm_model2 <- ca.jo(df_final_ts, ecdet = "none", type = "trace", K=k_order, spec = "transitory", season = 7)

summary(vecm_model2)

```

### Forecasting with VECM model

#### Using first Cointegration point vs existing data

```{r}

require(vars)

### Converting back VECM to new VAR model

VAR_pos_R <- lapply(c(1:5), function(x) vec2var(vecm_model1, r=x))

VAR_pos_R_seas <- lapply(c(1:5), function(x) vec2var(vecm_model2, r=x))

#VAR_pos_R1 <- vec2var(vecm_model1, r=1)
#VAR_pos_R2 <- vec2var(vecm_model1, r=2)
#VAR_pos_R3 <- vec2var(vecm_model1, r=3)
#VAR_pos_R4 <- vec2var(vecm_model1, r=4)
#VAR_pos_R5 <- vec2var(vecm_model1, r=5)


###########################################
#                                         #
### Prediction using the VECM_TO_VAR_R1 ###

### Prediction with test dataset

ndays <- ndays1

VAR_pos_R_pred <- lapply(VAR_pos_R, function(x) predict(x, new_data=df_test_ts, n.ahead=ndays))

### Prediction with Seasonality
VAR_pos_R_seas_pred <- lapply(VAR_pos_R_seas, function(x) predict(x, new_data=df_test_ts, n.ahead=ndays))


### Plotting predictions

plt_R <- lapply(VAR_pos_R, function(x) var_pred(model = x, df=df_test_ts, ahead=ndays))

plt_R_seas <- lapply(VAR_pos_R_seas, function(x) var_pred(model = x, df=df_test_ts, ahead=ndays))


#var_pred(model = VAR_pos_R2,  df=df_test_ts, ahead=14)

#var_pred(model = VAR_pos_R3,  df=df_test_ts, ahead=14)

#var_pred(model = VAR_pos_R4,  df=df_test_ts, ahead=14)

#var_pred(model = VAR_pos_R5,  df=df_test_ts, ahead=14)

### Forecasting auto-regressively

#var_forecast(model = VAR_pos_R1, ahead=14, actual_df=df_final)

## Uncomment
#var_forecast(model = VAR_pos_R[[1]], ahead=14, actual_df=df_final)
#var_forecast(model = VAR_pos_R[[2]], ahead=14, actual_df=df_final)
#var_forecast(model = VAR_pos_R[[3]], ahead=14, actual_df=df_final)
#var_forecast(model = VAR_pos_R[[4]], ahead=14, actual_df=df_final)
#var_forecast(model = VAR_pos_R[[5]], ahead=14, actual_df=df_final)

save(VAR_pos_R, file = "VAR_pos_R.rda")
save(VAR_pos_R_seas, file = "VAR_pos_R_seas.rda")
#save(VAR_pos_R_pred, file = "VAR_pos_R_pred.rda")
#save(VAR_pos_R_seas_pred, file = "VAR_pos_R_seas_pred.rda")
#save(plt_R, file = "plt_R.rda")
#save(plt_R_seas, file = "plt_R_seas.rda")

plt_R_seas

################### The End #####################

```


<p><p>


## USING THE Full dataset FOR FORECASTING FUTURE DATES WITHOUT AVAILABLE PREDICTORS DATA ##

```{r}

##### Test with full data

#Lag selection
var_lag1 <- VARselect(df_full_ts, lag.max = 8, type = "both", season = 7)

# Extracting the lag seleciton by AIC
max_lag1 = as.numeric(var_lag1$selection[1])

## Option 1: VAR Model
VAR_pos1 <- VAR(df_full_ts, lag.max = max_lag, type = "both", season = 7)

summary(VAR_pos1$varresult$pos_ma)


```
```{r}

k_order1 <- VAR_pos1$p

vecm_model11 <- ca.jo(df_full_ts, ecdet = "none", type = "trace", K=k_order, spec = "transitory")

vecm_model21 <- ca.jo(df_full_ts, ecdet = "none", type = "trace", K=k_order, spec = "transitory", season = 7)

summary(vecm_model21)


#vecm_final <- cajolst(vecm_model11)

```


```{r}

VAR_pos_Rs <- lapply(c(1:5), function(x) vec2var(vecm_model11, r=x))
save(VAR_pos_Rs, file="VAR_pos_Rs.rda")

VAR_pos_Rs_seas <- lapply(c(1:5), function(x) vec2var(vecm_model21, r=x))
save(VAR_pos_Rs_seas, file="VAR_pos_Rs_seas.rda")


VAR_pos_Rs_pred <- lapply(VAR_pos_Rs, function(x) predict(x,  n.ahead=ndays))

### Prediction with Seasonality
VAR_pos_Rs_seas_pred <- lapply(VAR_pos_Rs_seas, function(x) predict(x, new_data=df_full_ts, n.ahead=ndays))


### Plotting predictions

plt_Rs <- lapply(VAR_pos_Rs, function(x) var_forecast(model = x, ahead=ndays))

plt_Rs_seas <- lapply(VAR_pos_Rs_seas, function(x) var_pred(model = x, ahead=ndays))


plt_Rs

```
### VECM model selection based on MAPE


```{r}

require("ie2misc")

mape_RS <- lapply(1:length(VAR_pos_Rs), function(x) mape(VAR_pos_Rs_pred[[x]]$fcst$pos_ma[,"fcst"], df_test[1:ndays,"pos_ma"]))


mape_Seas_RS <- lapply(1:length(VAR_pos_Rs_seas), function(x) mape(VAR_pos_Rs_seas_pred[[x]]$fcst$pos_ma[,"fcst"], df_test[1:ndays,"pos_ma"]))

## Extracting the position of the lowest 
a <- which(unlist(mape_RS)==min(unlist(mape_RS)))

## Extracting the position of the lowest 
b <- which(unlist(mape_Seas_RS)==min(unlist(mape_Seas_RS)))



### Moving the old model to the old folder before saving the new model

if(file.exists("models/final_model.rda")){
    
  file.rename("models/final_model.rda", paste0("models/",today(),"_final_model.rda"))
}



## Selecting the minimal MAPE from seasonal and non-seasonal
if(a < b){
  
  VAR_forecast <- predict(VAR_pos_Rs[[a]], n.ahead=ndays)
  print("The selected model is Without seasonality")
  final_model <-VAR_pos_Rs[[a]]
  
}else{
  VAR_forecast <- predict(VAR_pos_Rs_seas[[b]], n.ahead=ndays)
  print("The selected model is with seasonality")
  final_model <- VAR_pos_Rs_seas[[b]]
}

### Saving the model

save(final_model, file = "models/final_model.rda")


### Plotting Forecast
plot(VAR_forecast)

```


### Plotting entire dataset based on model selection

```{r}

### Creating the vector of days ahead corresponding to the forecast
days_ahead <- last(df_final$date)+days(0:ndays)[-1]
fcst_ahead <- VAR_forecast$fcst$pos_ma[,"fcst"]

### only the data part of the forecast ndays ahead
df_fcst_only <- tibble(date=days_ahead, 
                       value=VAR_forecast$fcst$pos_ma[,"fcst"], 
                       lower=VAR_forecast$fcst$pos_ma[,"lower"],
                       upper=VAR_forecast$fcst$pos_ma[,"upper"])

### Binding the dates of the actual to the days ahead and the Actual values to the prediction ahead
df_fcst <- tibble(date=c(df_final$date, days_ahead), value=c(df_final$pos_ma,VAR_forecast$fcst$pos_ma[,"fcst"]))

plot_ly() %>%
  add_lines(data = tail(df_fcst,60), x=~date, y=~value, name="Positivity rate") %>%
  add_ribbons(data = df_fcst_only, 
              x=~date,
              ymin = ~lower,
              ymax = ~upper,
              color=I("orange"),
              name = "95% confidence")

```

### Predicting Zipcode level data using ADI-Cases Coefficients
##### ADI_Cases rate will be used to estimate positivity rate using a 2SLS (2 Stages least square regression)
##### First stage being the regression between ADI and Case rates in Harris
##### Second being regression between ADI and positivity rate in Tarrant
##### Tarrant county zipcode with median ADI is assigned the lowest 

```{r}
## Uncomment before running if var
load("data/harris_lm.rda")
load("data/tarrant_data.rda")
adi_case_coef <- as.numeric(harris_lm$coefficients[1])

```


#### Computing positivity rate zipcode level

```{r}

### Selecting reference zipcode
ref_zip <- tarrant_data[round(tarrant_data$adi_st_mean,1)==round(median(tarrant_data$adi_st_mean),1),"zipcode"]

ref_adi <- tarrant_data$adi_st_mean[tarrant_data$zipcode==ref_zip]

### Calculating the adi difference from the reference and the Additive change in positivity rate based on 
### Calculated adi-case_rate coefficient
tarrant_data <- tarrant_data %>% dplyr::mutate(adi_diff=adi_st_mean-ref_adi, change=adi_diff*adi_case_coef)


### Creating a matrix of zipcodes
tarrant_data_zip <- spread(tarrant_data, zipcode, change) %>% dplyr::select(-c(1:6))

### Expanding dataframe with NA values for merging purpose
tarrant_data_zip[nrow(tarrant_data)+nrow(df_fcst)-nrow(tarrant_data),] <- NA

### Adding the county positivity rate in front

tarrant_data_zip <- cbind(df_fcst, tarrant_data_zip)

### Assigning default to reference zip

#tarrant_data_zip[,ref_zip] <- tarrant_data_zip$value

### Assigning Tarrant County positivity rate including predicted to the reference zipcode

tarrant_zip_est <- lapply(names(tarrant_data_zip)[-c(1:2)], function(x) tarrant_data_zip$value+mean(tarrant_data_zip[,x], na.rm=T))

### Binding columns of the dataframe
tarrant_zip_est <- bind_cols(tarrant_zip_est)

### Reassigning zipcode names to the list

names(tarrant_zip_est)<- paste0("Z", names(tarrant_data_zip[-c(1:2)]))

View(tarrant_zip_est)
tarrant_zip_est <- bind_cols(df_fcst, tarrant_zip_est)
tarrant_zip_est <- tarrant_zip_est %>% mutate_if(is.numeric, round, digits=2)
save(tarrant_zip_est, file = "data/tarrant_zip_est.rda")


```





```{r}

plot_ly(data=tarrant_zip_est, x=~date) %>% 
  add_trace(y=~Z76105, mode="markers+lines", name="76105") %>% 
  add_trace(y=~Z76106, mode="markers+lines", name="76106") %>%
  add_trace(y=~Z76164, mode="markers+lines", name="76164")

```


## Forecast


### VAR Forecast with newdata

```{r}

library(tsDyn)

vecm1 <- VECM(df_final_ts, lag = 8)

vecm1 %>% summary()

VARrep(vecm1)

vecm1_pred <- predict_rolling(vecm1, newdata = df_test ,nroll = 7)

plot(vecm1_pred$true$pos_ma, vecm1_pred$pred$pos_ma, type="p")

```

